Microbiome analysis - machine learning

### 2020-09-14 code/R_markdown/P1_analysis_prep_figures.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_plot_2020_09_02.RData")

@@

# Generate supplemental file 1B to get read counts

@@

# Plot positive controls

@@

# Figure 2

@@

# Table 1
```{r}

@@

>>>>

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_plot_2020_09_11.RData")

@@

# Generate supplemental file 1D to get read counts

@@

# Plot positive controls (Supplemental Figure 1A-B)

@@

# Supplemental Figure 4

@@

# Table 1

@@

How many ASVs map to Prevotella? (For Fig 3 caption)
```{r}

ntaxa(subset_taxa(ps_rare_diff, Genus == "Prevotella"))
ntaxa(subset_taxa(ps_rare_diff, Genus == "Bifidobacterium"))

ntaxa(subset_taxa(ps_rare_diff, Genus == "Megasphaera"))
ntaxa(subset_taxa(ps, Genus == "Megasphaera"))

### 2020-09-14 code/R_markdown/P2_cohort.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_02.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")

>>>>

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")

### 2020-09-14 code/R_markdown/P2_cohort_figures.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load packages
library(ggplot2); library(rmarkdown); library(vegan); library(dotwhisker)
library(dplyr); library(phyloseq)

@@

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P2output_plot_2020_09_02.RData")

@@

>>>>

# load packages
library(ggplot2); library(rmarkdown); library(vegan); library(dotwhisker)
library(dplyr); library(phyloseq); library(tibble)

@@

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P2output_plot_2020_09_11.RData")

@@

ps_rare_for_beta_mode_sample_data = ps_rare_for_beta
qu = quantile(sample_data(ps_rare_for_beta_mode_sample_data)$Age.in.days, c(0.25, 0.5, 0.75))
temp_sample_data = as(sample_data(ps_rare_for_beta_mode_sample_data),"data.frame") %>% 
  rownames_to_column('sample') %>%
  mutate(age_quartile = as.factor(ifelse(Age.in.days < qu[[1]][1], 1,
                               ifelse(Age.in.days > qu[[3]][1], 4,
                                      ifelse(Age.in.days > qu[[2]][1], 3,2)))),
         hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)),
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>%
    column_to_rownames('sample')
sample_data(ps_rare_for_beta_mode_sample_data) = temp_sample_data
p = plot_ordination(
  physeq = ps_rare_for_beta_mode_sample_data,
  ordination = ps_rare.ord,
  color = "Clinical.type",
  shape = "age_quartile"
)   + 
  scale_color_manual(values = c("#666666","#FF7F00","#1F78B4"),
                     labels = c("12" = "diarrheal","10" = "subclinical","1" = "pre-detection") )+ # 12 is diarrhea, 1,2 are pre
  #scale_shape_manual(values = c(5,16,4),
   #                  labels = c("1" = "youngest quartile","2" = "middle","4" = "oldest quartile") )+
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))  + 
  labs(color = "symptomatic?", shape = "Age (quartitled)")
p

p = plot_ordination(
  physeq = ps_rare_for_beta_mode_sample_data,
  ordination = ps_rare.ord,
  color = "Clinical.type",
  shape = "meleagridis"
)   + 
  scale_color_manual(values = c("#666666","#FF7F00","#1F78B4"),
                     labels = c("12" = "diarrheal","10" = "subclinical","1" = "pre-detection") )+ # 12 is diarrhea, 1,2 are pre
  scale_shape_manual(values = c(5,16),
                     labels = c("1" = "positive","0" = "negative") )+
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))  + 
  labs(color = "symptomatic?", shape = "C. meleagridis")
p

p = plot_ordination(
  physeq = ps_rare_for_beta_mode_sample_data,
  ordination = ps_rare.ord,
  color = "Clinical.type",
  shape = "hominis"
)   + 
  scale_color_manual(values = c("#666666","#FF7F00","#1F78B4"),
                     labels = c("12" = "diarrheal","10" = "subclinical","1" = "pre-detection") )+ # 12 is diarrhea, 1,2 are pre
  scale_shape_manual(values = c(5,16),
                     labels = c("1" = "positive","0" = "negative") )+
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"))  + 
  labs(color = "symptomatic?", shape = "C. hominis")
p

### 2020-09-14 code/R_markdown/P3_machine_learning.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
total_diarrhea = total_diarrhea  %>% filter(CHILDID %in% burden_mirpur$SID) # remaining kids never were sick
all_cause_diar = total_diarrhea %>% group_by(CHILDID,study) %>% summarize(episodes = n()) %>% ungroup()
rm(list= ls()[! (ls() %in% c('all_cause_diar','SEED_INT','colors'))])
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_02.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P2output_2020_09_02.RData")

@@

x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
rf <- randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.00015,130))
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_PD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))   
x_for_PD = x
y_for_PD = y

extracted_importance = importance(rf,type=2) # type 2 = node impurity, more useful for auc
extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
top10_PD = tail(extracted_importance,10)

@@

compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
x_for_TOD = x
y_for_TOD = y

rf = randomForest_res(x_for_TOD,y_for_TOD,10)
top10_TOD = rf[[4]]
roc_TOD = rf[[2]]
AUC_TOD = rf[[3]]
AUC_TOD

@@

# time-of-detection
compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.numeric(as.character(sample_data(ps_for_class)$Number.of.crypto.positive.events))
y[y!=1] = 2 # merge third infections with second

@@

# predetection
compare <- c(1,2) # PD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y <- as.numeric(as.character(sample_data(ps_for_class)$Number.of.crypto.positive.events))
y[y!=1] = 2 # merge third infections with second
rf = randomForest(x,as.factor(y),importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.1,100))
rf
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_PD_1stv2nd_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

@@

compare <- c(10,12) # TOD +/- symptoms
ps_for_class2 <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
all_cause_diar_temp = all_cause_diar %>% filter(CHILDID %in% sample_data(ps_for_class2)$Child.ID)
all_cause_diar_temp = merge(all_cause_diar_temp, as(sample_data(ps_for_class2)[,1:2], "data.frame"), by.x = "CHILDID", by.y = "Child.ID", all.y = T) %>% select(-study,-CHILDID)
sample_data(ps_for_class2)$episodes = all_cause_diar_temp$episodes[order(match(all_cause_diar_temp$sample_id_mod, sample_data(ps_for_class2)$sample_id_mod))]
x <- cbind(otu_table(ps_for_class2), as(sample_data(ps_for_class2),"data.frame") %>% select(Age.in.days))
y <- sample_data(ps_for_class2)$episodes
## WHY IS THERE AN NA
y[is.na(y)] = 0

@@

compare <- c(1,2) # PD +/- symptoms
ps_for_class2 <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
all_cause_diar_temp = all_cause_diar %>% filter(CHILDID %in% sample_data(ps_for_class2)$Child.ID)
all_cause_diar_temp = merge(all_cause_diar_temp, as(sample_data(ps_for_class2)[,1:2], "data.frame"), by.x = "CHILDID", by.y = "Child.ID", all.y = T) %>% select(-study,-CHILDID)
sample_data(ps_for_class2)$episodes = all_cause_diar_temp$episodes[order(match(all_cause_diar_temp$sample_id_mod, sample_data(ps_for_class2)$sample_id_mod))]
x <- cbind(otu_table(ps_for_class2), as(sample_data(ps_for_class2),"data.frame") %>% select(Age.in.days))
y <- sample_data(ps_for_class2)$episodes
## WHY IS THERE AN NA
y[is.na(y)] = 0

@@

x <- cbind(cbind(time_of_detect_microbes,surv_microbes), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days))
y = as.factor(sample_data(subset_samples(ps_for_class, Clinical.type %in% compare))$Clinical.type)
rf = randomForest_res(x,y,15)
rf[[1]]

@@

x_withabx = cbind(cbind(abx_history, otu_table(for_metadata)), as(sample_data(for_metadata),"data.frame") %>% select(Age.in.days))
y = as.factor(sample_data(for_metadata)$Clinical.type)

rf = randomForest_res(x_withabx,y,25)
rf[[1]]

@@

plot_tax = tax2 %>% select(pre.detection, time.of.detection, var) %>% melt(id.vars = "var", na.rm = T)
plot_tax$var = as.factor(plot_tax$var)
plot_tax$var = reorder(plot_tax$var, -plot_tax$value)
plot_tax$var <- relevel(plot_tax$var, "ASV19 (Megasphaera)")
plot_tax_4D = plot_tax

@@

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
#ASV10 (Collinsella) for PD
#ASV5 (Parabacteroides) for TOD
TOD_twoASVs = tax2 %>% filter(var == 'ASV10 (Megasphaera)' | var == 'ASV5 (Parabacteroides)')

@@

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
#ASV10 (Collinsella) for PD
#ASV5 (Parabacteroides) for TOD
PD_twoASVs = tax2 %>% filter(var == 'ASV10 (Megasphaera)' | var == 'ASV3 (Collinsella)')
tax2_PD = tax2

@@

twoASVs = rbind(TOD_twoASVs %>% filter(var == "ASV10 (Megasphaera)") %>% mutate(time = 'time-of-detection'), 
                PD_twoASVs %>% filter(var == "ASV10 (Megasphaera)") %>% mutate(time = 'pre-detection'))

@@

twoASVs_PD_wide = tidyr::spread(PD_twoASVs %>% select(value_use,symptomatic_use,sample, Genus),
                             key = Genus, value = value_use) %>% 
  mutate(`symptomatic?` = as.factor(symptomatic_use))

@@

plt_df1 = as.data.frame(top10_TOD) %>%rownames_to_column('variable') %>% dplyr::rename(value = top10_TOD) 
plt_df2 = as.data.frame(top20_TOD_with_abx) %>%rownames_to_column('variable') %>% dplyr::rename(value = top20_TOD_with_abx) %>% top_n(10,value)
plt_df = rbind(plt_df1 %>% mutate(model = 'without abx'),plt_df2 %>% mutate(model = 'with abx'))
plt_df$model = as.factor(plt_df$model)

@@

## DATA NOT SHOWN: Does microbiome predict Crypto species at TOD? HERE

>>>>

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
total_diarrhea = total_diarrhea  %>% filter(CHILDID %in% burden_mirpur$SID) # remaining kids never were sick
all_cause_diar = total_diarrhea %>% group_by(CHILDID,study) %>% summarize(episodes = n()) %>% ungroup()
rm(list= ls()[! (ls() %in% c('all_cause_diar','SEED_INT','colors'))])
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P2output_2020_09_11.RData")

@@

x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
rf <- randomForest(x,y,importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.00015,130))
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_PD_" , format(Sys.time(), "%Y_%m_%d")),".RData"))   
x_for_PD = x
y_for_PD = y

extracted_importance = importance(rf,type=2) # type 2 = node impurity, more useful for auc
extracted_importance = extracted_importance[rownames(extracted_importance)[order(extracted_importance)],]
top10_PD = tail(extracted_importance,15)

@@

compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- as.factor(sample_data(ps_for_class)$Clinical.type)
x_for_TOD = x
y_for_TOD = y

rf = randomForest_res(x_for_TOD,y_for_TOD,15)
top10_TOD = rf[[4]]
roc_TOD = rf[[2]]
AUC_TOD = rf[[3]]
AUC_TOD

@@

# time-of-detection
compare <- c(10,12) # TOD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- as.numeric(as.character(sample_data(ps_for_class)$Number.of.crypto.positive.events))
y[y!=1] = 2 # merge third infections with second

@@

# predetection
compare <- c(1,2) # PD +/- symptoms
ps_for_class <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
x <- cbind(otu_table(ps_for_class), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- as.numeric(as.character(sample_data(ps_for_class)$Number.of.crypto.positive.events))
y[y!=1] = 2 # merge third infections with second
rf = randomForest(x,as.factor(y),importance = TRUE, ntree = 50*dim(x)[[2]], replace=TRUE, nodesize=1, classwt=c(0.1,100))
rf
save(list = c("rf"), file = paste0(paste0("~/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_PD_1stv2nd_" , format(Sys.time(), "%Y_%m_%d")),".RData"))

@@

compare <- c(10,12) # TOD +/- symptoms
ps_for_class2 <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
all_cause_diar_temp = all_cause_diar %>% filter(CHILDID %in% sample_data(ps_for_class2)$Child.ID)
all_cause_diar_temp = merge(all_cause_diar_temp, as(sample_data(ps_for_class2)[,1:2], "data.frame"), by.x = "CHILDID", by.y = "Child.ID", all.y = T) %>% select(-study,-CHILDID)
sample_data(ps_for_class2)$episodes = all_cause_diar_temp$episodes[order(match(all_cause_diar_temp$sample_id_mod, sample_data(ps_for_class2)$sample_id_mod))]
x <- cbind(otu_table(ps_for_class2), as(sample_data(ps_for_class2),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- sample_data(ps_for_class2)$episodes
## WHY IS THERE AN NA
y[is.na(y)] = 0

@@

compare <- c(1,2) # PD +/- symptoms
ps_for_class2 <- subset_samples(ps_rare_diff, Clinical.type %in% compare)
all_cause_diar_temp = all_cause_diar %>% filter(CHILDID %in% sample_data(ps_for_class2)$Child.ID)
all_cause_diar_temp = merge(all_cause_diar_temp, as(sample_data(ps_for_class2)[,1:2], "data.frame"), by.x = "CHILDID", by.y = "Child.ID", all.y = T) %>% select(-study,-CHILDID)
sample_data(ps_for_class2)$episodes = all_cause_diar_temp$episodes[order(match(all_cause_diar_temp$sample_id_mod, sample_data(ps_for_class2)$sample_id_mod))]
x <- cbind(otu_table(ps_for_class2), as(sample_data(ps_for_class2),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y <- sample_data(ps_for_class2)$episodes
## WHY IS THERE AN NA
y[is.na(y)] = 0

@@

x <- cbind(cbind(time_of_detect_microbes,surv_microbes), as(sample_data(ps_for_class),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y = as.factor(sample_data(subset_samples(ps_for_class, Clinical.type %in% compare))$Clinical.type)
rf = randomForest_res(x,y,15)
rf[[1]]

@@

x_withabx = cbind(cbind(abx_history, otu_table(for_metadata)), as(sample_data(for_metadata),"data.frame") %>% select(Age.in.days, LIB13.C..hominis, C..mel.y)) %>% 
  mutate(hominis = as.factor(ifelse(LIB13.C..hominis <40,1,0)), 
         meleagridis = as.factor(ifelse(C..mel.y <40,1,0))) %>% select(-LIB13.C..hominis,-C..mel.y)
y = as.factor(sample_data(for_metadata)$Clinical.type)

rf = randomForest_res(x_withabx,y,15)
rf[[1]]

@@

plot_tax = tax2 %>% select(pre.detection, time.of.detection, var) %>% melt(id.vars = "var", na.rm = T)
plot_tax$var = as.factor(plot_tax$var)
plot_tax$var = reorder(plot_tax$var, -plot_tax$value)
plot_tax$var <- relevel(plot_tax$var, levels(plot_tax$var)[grep("Megasphaera", levels(plot_tax$var))])
plot_tax_4D = plot_tax

@@

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
#ASV10 (Collinsella) for PD
#ASV5 (Parabacteroides) for TOD
TOD_twoASVs = tax2 %>% filter(var ==  tax2$var[grep("Megasphaera", tax2$var)] | 
                                var == tax2$var[grep("Parabacteroides", tax2$var)])

@@

tax = tax_table(ps_rare_diff)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
#ASV10 (Collinsella) for PD
#ASV5 (Parabacteroides) for TOD
PD_twoASVs = tax2 %>% filter(var == tax2$var[grep("Megasphaera", tax2$var)] | 
                                var == tax2$var[grep("Collinsella", tax2$var)])
tax2_PD = tax2

@@

twoASVs = rbind(TOD_twoASVs %>% filter(var == TOD_twoASVs$var[grep("Megasphaera", TOD_twoASVs$var)]) %>% 
                  mutate(time = 'time-of-detection'), 
                PD_twoASVs %>% filter(var == PD_twoASVs$var[grep("Megasphaera", PD_twoASVs$var)]) %>% 
                  mutate(time = 'pre-detection'))

@@

twoASVs_PD_wide = tidyr::spread(PD_twoASVs %>% select(value_use,symptomatic_use,var, sample, Genus),
                             key = var, value = value_use) %>% 
  mutate(`symptomatic?` = as.factor(symptomatic_use))

@@

plt_df1 = as.data.frame(top10_TOD) %>%rownames_to_column('variable') %>% dplyr::rename(value = top10_TOD) 
plt_df2 = as.data.frame(top20_TOD_with_abx) %>%rownames_to_column('variable') %>% dplyr::rename(value = top20_TOD_with_abx)
plt_df = rbind(plt_df1 %>% mutate(model = 'without abx'),plt_df2 %>% mutate(model = 'with abx'))
plt_df$model = as.factor(plt_df$model)

@@

## DATA NOT SHOWN: Does microbiome predict Crypto species at TOD?
